---
title: "Payero and Sekaran review"
author: "C A Pellett"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
pivots0 <- readr::read_csv("SC_Pivots_2021_Lat_Lon2.csv")

```

The research by Payero and Sekaran resulted in a table of center pivots with `r nrow(pivots0)` rows and `r ncol(pivots0)` columns: `r names(pivots0)`.

Strange that the file name indicates the data is for 2021, but the MapYear column indicates 2022.

# id Column
```{r}
scutils::dupes(pivots0, 'id') # %>% View()
## There are many duplicate id's.

scutils::dupes(pivots0, c('Long', 'Lat'))
## But there aren't any duplicate locations.
```
There are many duplicate values in the id column, but there aren't any duplicate locations according to the Lat and Long columns.

# Length columns
```{r}
ggplot2::qplot(pivots0$length_ft, pivots0$WetRad_ft)

```
This looks about right. The wetted radius is almost always equal or greater than the pivot length, because end guns irrigate beyond the physical footprint of the pivot.

# The Number of Towers
```{r}
ggplot2::qplot(pivots0$NumTowers)

```
Looks about right to me.

# The Degrees of rotation
```{r}
ggplot2::qplot(pivots0$Degrees)

```
Looks about right to me.

# Comments
```{r}
dplyr::filter(pivots0, !is.na(Comment))

```
For some reason the comments say "Corner system", but the Degrees is 360. Strange. I guess corner system doesn't mean what I think it means.

# CountyName Column
```{r}
unique(pivots0$CountyName)
## todo: test this with a spatial join.
```
Looks right. I could test this with a spatial join.

# Spatial Coordinates
```{r}
ggplot2::qplot(pivots0$Long, pivots0$Lat)

```
Looks about right to me.

## MapYear 
```{r}
unique(pivots0$MapYear)

```
The only value in this column is 2022. Yet, the file name indicates 2021.


```{r}
pivot_pts2022 <- pivots0 %>%
  dplyr::select(-id, -MapYear) %>%
  dplyr::mutate(
    PointID = 1:n(),
    length_ft = units::set_units(length_ft, 'ft'),
    WetRad_ft = units::set_units(WetRad_ft, 'ft')) %>%
  sf::st_as_sf(coords=c('Long', 'Lat'), crs=4326)

```

```{r}
## Calculate Area (acres)
## st_area returns m2 for this CRS.

pivot_circles2022 <- pivot_pts2022 %>%
  sf::st_buffer(., dist=.$length_ft, nQuadSegs=6) %>%
  dplyr::mutate(Acres = units::drop_units(
    sf::st_area(geometry)*(Degrees/360)/4046.86)) 

pivot_wetted2022 <- pivot_pts2022 %>%
  sf::st_buffer(., .$WetRad_ft, nQuadSegs=6) %>%
  dplyr::mutate(Acres = units::drop_units(
    sf::st_area(geometry)*(Degrees/360)/4046.86))

```


```{r}
usethis::use_data(pivot_pts2022, overwrite=T)

usethis::use_data(pivot_circles2022, overwrite=T)

usethis::use_data(pivot_wetted2022, overwrite=T)


## write shapefiles or geoJSON or whatever.

```


