---
title: "03_combined.Rmd"
author: "C A Pellett"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(units)
```

```{r eval=F}
# sf::st_write(irrigation_zones2020, 'irrigation_zones2020.shp')
# sf::st_write(pivot_circles2022, 'pivot_circles2022.shp')
```

# Spatial overlay 
Spatial joins are used to combine the two editions of data. The result is a new column for each dataset which holds an identifier key for each irrigation distribution system.


```{r}
load("~/Rpackages/sccenterpivot/data/pivot_pts2020.rda")
load("~/Rpackages/sccenterpivot/data/pivot_pts2022.rda")

pts.a0 <- pivot_pts2020 %>%
  dplyr::rename(id.a = PointID) %>%
    sf::st_transform(32133) ## SC State Plane NAD83


pts.b0 <- pivot_pts2022 %>%
  dplyr::rename(id.b = PointID) %>%
  sf::st_transform(32133) ## SC State Plane NAD83 
```

## Combine the pivot points.

```{r}

nearest_pts.a <- sf::st_nearest_feature(pts.a0, pts.b0)

distance_pts.a <- sf::st_distance(pts.a0, pts.b0[nearest_pts.a,], by_element = TRUE)

pts.a1 <- pts.a0 %>%
  dplyr::mutate(
    nearest.b = pts.b0[nearest_pts.a, 'id.b', drop=T],
    distance.b = distance_pts.a)

## ggplot2::qplot(distance_pts.a)
#### like 90% are less than 10 meters, so that is good. 
#### the very high values should probably be dropped.

```


```{r}
nearest_pts.b <- sf::st_nearest_feature(pts.b0, pts.a0)

distance_pts.b <- sf::st_distance(pts.b0, pts.a0[nearest_pts.b,], by_element = TRUE)

pts.b1 <- pts.b0 %>%
  dplyr::mutate(
    nearest.a = pts.a0[nearest_pts.b, 'id.a', drop=T],
    distance.a = distance_pts.b)

```

```{r}
joinsa.0 <- pts.a1[, c('id.a', 'nearest.b', 'distance.b'), drop=T]

joinsa.1 <- joinsa.0 %>%
  dplyr::group_by(nearest.b) %>%
  dplyr::filter(distance.b == min(distance.b)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(id.a = as.character(id.a),
                nearest.b = as.character(nearest.b))

joinsb.0 <- pts.b1[, c('id.b', 'nearest.a', 'distance.a', 'length_ft'), drop=T] 

joinsb.1 <- joinsb.0 %>%
  dplyr::group_by(nearest.a) %>%
  dplyr::filter(distance.a == min(distance.a)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(id.b = as.character(id.b),
                nearest.a = as.character(nearest.a))

joinsb.2 <- joinsb.1 %>%
  dplyr::filter(distance.a < length_ft)
## this removes like 30 joins.

joins_perfect_pt <- joinsb.2 %>%
  dplyr::inner_join(joinsa.1, by=c('id.b'='nearest.b', 'nearest.a'='id.a')) %>%
  dplyr::filter(distance.a == distance.b) %>%
  dplyr::select(id.a=nearest.a, id.b, distance=distance.a) %>%
  dplyr::left_join(
    dplyr::select(pivot_pts2020, id.a=PointID, ZoneID) %>%
      dplyr::mutate(id.a = as.character(id.a)),
    'id.a')
## Convert these to lines.

```

If two points are nearest to each other, and the distance between them is less than the length_ft, and the pt.b is within the irrigation zone, then you can probably figure they are the same point.

```{r}

usethis::use_data(joins_perfect_pt, overwrite=T)

```

## Combine the Zones
```{r}
load("~/Rpackages/sccenterpivot/data/irrigation_zones2020.rda")
load("~/Rpackages/sccenterpivot/data/pivot_circles2022.rda")
```

```{r}
zones.a <- irrigation_zones2020 %>%
  dplyr::rename(id.a = ZoneID) %>%
    # sf::st_transform(32133)  %>% ## SC State Plane NAD83
  sf::st_make_valid()

zones.b <- pivot_circles2022 %>%
  dplyr::rename(id.b = PointID) %>%
    # sf::st_transform(32133) %>%
  sf::st_make_valid()
```
The two datasets of mapped irrigation (Pellett 2020, and Sekaran and Payero 2022) are combined to allow for analysis of change over time. While the first dataset (Pellett 2020) includes approximate polygons of irrigated area, the second dataset (Sekaran and Payero 2022) includes point locations, pivot lengths, and pivot degrees. The point locations have been buffered by the pivot lengths to create circular polygons, and acreage irrigated is estimated using the pivot degrees. Because there is no information on the bearing of the semi-circular pivots, the circular polygons which represent semi-circular pivots require additional editing.

Various geographic intersections are developed in order to properly combine the two datasets.

### Self-Intersection
```{r eval=F}
## This code is included in 01_Pellett_2020_prep.Rmd

# overlaps.a <- zones.a %>%
#   sf::st_intersection() %>%
#   dplyr::filter(n.overlaps > 1)  %>%
#   sf::st_collection_extract('POLYGON') %>%
#   dplyr::mutate(Acres0 = units::drop_units(
#     sf::st_area(geometry)/4046.86)) %>% 
#   dplyr::filter(Acres0 > 0) %>%
#   tidyr::unnest_wider(origins) %>%
#   dplyr::select(-...1) %>%
#   dplyr::rename(ZoneID1=ZoneID, ZoneID2=...2)

## The result is 24 polygons, each less than 5 sqft

```


```{r}
## There are some pivots that lie entirely within other pivots.
within.b <- zones.b %>%
  dplyr::mutate(within = sf::st_within(geometry),
                length.within = lengths(within)) %>%
  dplyr::filter(length.within > 1)

## I can drop them from this analysis.
zones.b1 <- zones.b %>% 
  dplyr::filter(!(id.b %in% within.b$id.b))
## I did this, but I shouldn't have.
## smaller pivots aren't necessarily actually located within the larger pivot area.
## remember, these polygons are circular buffers of points, 
## not accurate representations of the pivot's range of motion.
## Hopefully, I spotted any relevant smaller neighboring pivots,
## and included them in unjoined_a.shp


overlaps.b <- zones.b1 %>%
  sf::st_intersection() %>%
  dplyr::filter(n.overlaps > 1)  
## 942 self-intersections.

```

```{r eval=F}
## This should select the POLYGONS from multifeatures.
overlaps.b1 <- overlaps.b %>%
  sf::st_collection_extract('POLYGON') 


overlaps.b2 <- overlaps.b1 %>%
  dplyr::mutate(
    Acres0 = units::drop_units(
      sf::st_area(geometry)/4046.86),
    n = lengths(origins)) %>%
  dplyr::filter(Acres0 > 0) 

unique(overlaps.b2$n)
```
Some zones overlap other zones in the same dataset. This could confound subsequent steps of the analysis.

## Intersections between the two datasets
```{r eval=F}
# zones.c0 <- sf::st_intersection(zones.a, zones.b1)
# saveRDS(zones.c0, 'zones_c0.rds')

## This sometimes (but not always??) takes a really long time to run.
```

```{r}
zones.c0 <- readRDS('zones_c0.rds')

```

```{r}
zones.c1 <- sf::st_make_valid(zones.c0)

```
The 2020 zones are already delineated. Many of the 2022 zones are also in the 2020 dataset. Identification of overlap between the two datasets will allow for many of the 2020 delineations to be carried forward to 2022 without modification, or with minimal modification. 

I will borrow terminology from the USGS NHD data preparation and call this process "conflation." 

The intersecting areas, overlaps, will be handled in a multi-step process. We will start with the simplest case: areas which have 0 intersections in the other dataset. Then we will address areas with a single intersection (a one-to-one spatial join) with the other dataset. Finally we will review areas with multiple intersections between the two datasets.

### Zones with zero intersections
If a zone touches 0 zones from the other dataset, and does not correspond to a joined point, then that is an unjoined zone.

```{r}
unjoined.a0 <- zones.a %>%
  dplyr::filter(!(id.a %in% zones.c0$id.a))
## 277 entries

unjoined.a1 <- unjoined.a0 %>%
  dplyr::filter(!(id.a %in% joins_perfect_pt$ZoneID))
## still 277 entries, good.
## So, there are 2,698 - 277 = 2,421 zones in 2020 that might match a zone in 2022.

unjoined.b0 <- zones.b %>%
  dplyr::filter(!(id.b %in% zones.c0$id.b))
## 501 entries

unjoined.b1 <- unjoined.b0 %>%
  dplyr::filter(!(id.b %in% joins_perfect_pt$id.b))
## 494 entries. So, 7 entries are joined to a point but not a zone... hm...
```
In the 2020 dataset, there are 277 features that don't touch any feature in the 2022 dataset. None of those 277 features has a "perfect point join" as described above.

In the 2022 dataset, there are 501 features that don't touch any feature in the 2020 dataset. 7 of them have a "perfect point join," and they need to be reviewed.


```{r eval=F}
# ## Those 7 entries need to be reviewed.
# unjoined.b0 %>% 
#   dplyr::filter(id.b %in% joins_perfect_pt$id.b) %>%
#   sf::st_write('unjoined_b0_pt_join.shp')

### I did this part after I had already done some of the steps below.
### all 7 of these entries got delineated and added to the final dataset.
### So, I'm not sure I need to deal with it.

## There are 2,980 - 494 = 2,486 zones in 2022 that might match a zone in 2020.
```
Export to QGIS and inspect each unjoined zone with imagery. If a 2020 zone is absent from imagery, flag that by deleting the feature from the shapefile. Import the shapefile back to R. Assign year_2020 values to unjoined.a by joining back to shapefile.

Unjoined.b geometries are edited to reflect the partial circles.

```{r eval=F}
### I've already run this chunk, and then I edited the shapefiles.
### So don't try to run it again, or you might overwrite the shapefiles I edited.

# sf::st_write(unjoined.a1, 'unjoined_a1.shp')
## I deleted several shapes which no longer exist. 
## I also added some, and I may have modified a few.

# sf::st_write(unjoined.b1, 'unjoined_b1.shp')
## Double Check these in QGIS.
## I may have deleted a few that didn't seem active (solar panels)

```

TODO: Read in unjoined.a1. 
Geometries that are deleted from unjoined.a1 are 2020 & 2021 FALSE.
New zones in unjoined.a1 are 2021 TRUE.
(Filter out Georgia zones)
Edited geometries in unjoined.a1 take precedent.

Read in unjoined.b1.
Geometries that are deleted are 2021 FALSE.
Edited geometries take precedent.

### One-to-one Intersections
If a single zone from A and a single zone from B intersect with each other and not with any other zone, and the estimated acreages are close, and the overlap is significant, then that would be a perfect zonal join.
```{r}

sf::sf_use_s2(FALSE)
## The code chunk below doesn't run unless I use the command above first.

joins_zone0 <- zones.c0 %>%
  dplyr::filter(
    !scutils::duplicated2(id.a) & 
      !scutils::duplicated2(id.b)) %>%
  dplyr::mutate(
    Acres_difference = abs(Acres-Acres.1),
    Acres_average = (Acres + Acres.1)/2,
    Acres_overlap = units::drop_units(sf::st_area(geometry)/4046.86),
    Acres_percent_difference = Acres_difference/Acres_average,
    Overlap_percent = 1 - ((Acres - Acres_overlap)/Acres),
    id.b=as.character(id.b),
    join_score = 1 - ((1-Acres_percent_difference)*Overlap_percent)) %>%
  dplyr::rename(ZoneID = id.a) %>%
  dplyr::left_join(
    joins_perfect_pt %>%
      sf::st_drop_geometry() %>%
      dplyr::mutate(distance_m = units::drop_units(distance)) %>%
      dplyr::select(ZoneID, id.b, distance) %>%
      dplyr::filter(!is.na(ZoneID)),
    by=c('ZoneID', 'id.b'))

### If I don't set S2 to False, I get this error:
## ERROR: Edge 27 is degenerate (duplicate vertex)

## note: Acres_overlap compares the full circle of the 2022 buffered points.
```
I create a Join Rank score by multiplying the percent difference in the calculated acreages of the two zones times the inverse of the percent of overlap. So, two zones that have exactly the same acreage and overlap perfectly would have a join rank score of 1. Export these features to QGIS. Inspect them, starting at the lowest join rank scores. Once they start looking good, can stop inspecting. The zones A geometries will take precedent.

```{r}

usethis::use_data(joins_zone0)
### I want to calculate a summary error statistic for Pellett 2023a JSCWRC publication.
```




```{r}
# ggplot2::qplot(
#   x=joins_zone0$Acres_percent_difference, 
#   y=joins_zone0$Overlap_percent)
# 
# ggplot2::qplot(joins_zone0$join_score)

# joins_perfect_zone <- joins_zone0 %>%
#   dplyr::filter(Acres_percent_difference < 0.2)

```


```{r}
joined.a0 <- zones.a %>%
  dplyr::inner_join(
    joins_zone0 %>%
      dplyr::select(id.a=ZoneID, join_score) %>%
      sf::st_drop_geometry(),
    by='id.a')

```

Review the joined.a polygons. Edit geometry as needed.
```{r eval=F}
## sf::st_write(joined.a0, 'joined_a0.shp')
### I've already run this, and then edited the shapefile.
### so don't overwrite the shapefile.
```
I went through and verified each one. The Join_rank score didn't seem to work so well. Hard to tell for sure. Whenever I saw a new center pivot, I added it to unjoined.a.shp.


### One to Many, Many to One, and Many to Many Intersections
```{r}
multi_joins_zone0 <- zones.c0 %>%
  dplyr::filter(
    scutils::duplicated2(id.a) |
      scutils::duplicated2(id.b))

multi_joins_zone1.a <- multi_joins_zone0 %>%
  sf::st_drop_geometry() %>%
  dplyr::semi_join(zones.a, ., 'id.a')

multi_joins_zone1.b <- multi_joins_zone0 %>%
  sf::st_drop_geometry() %>%
  dplyr::semi_join(zones.b1, ., 'id.b')
```


```{r}
### This is what I'd like to do:
## get the zones.a and zones.b that are in this set.
## spatial dissolve the pivots into larger polygons.
## then, in qgis, iterate through the merged larger polygons,
## while editing the zones.a polygons. (joined.a1)

## then bring the zones.a polygons back in to R.

### None of these things work:
# multi_joins_zones1 <- sf::st_combine(multi_joins_zone1.a, multi_joins_zone1.b)
# multi_joins_zones2 <- sf::st_union(multi_joins_zone1.a, multi_joins_zone1.b, by_feature=T)
# multi_joins_zones2 <- sf::st_union(multi_joins_zones1, )

multi_joins_zone2 <- dplyr::bind_rows(
   dplyr::select(multi_joins_zone1.a, geometry),
   dplyr::select(multi_joins_zone1.b, geometry))

### This doesn't work either:
# multi_joins_zones3 <- sf::st_union(multi_joins_zone2, by_feature=T)

### (by "doesn't work" I mean it doesn't union like I would expect)
### apparently sf doesn't have native implementation of union as i would expect.
### igraphs and terra packages can be used to merge clusters.
### but it is kinda complicated and unnecessary.
```


```{r eval=F}
# sf::st_write(multi_joins_zone2, 'multi_joins_zone2.shp')

### just union multi_joins_zone2 in qgis.
## I used the qgis dissolve tool, 
## which turned them all into a single multi-part polygon feature.
## then I used qgis multipart to singlepart tool.
## that reduces the number of features from 2,422 to 461
## So I can iterate through them and inspect 
## without looking at each set of overlapping polygons 
## multiple times.

### And then, iterate through the unioned zones created as described above,
## To efficiently review these zones:
# sf::st_write(multi_joins_zone1.a, 'multi_joins_zone_1a.shp')

### I edited the multi_joins_zone1.a shapefile in QGIS, so don't overwrite it.
```


```{r}
## combine the unjoined_a1.shp, unjoined_b.shp, joined_a0.shp, and multi_joins_a1.shp

## merge them wisely to conserve the year values.

### Start with unjoined.b: this is new pivots that Sekaran and Payero found.
### Go ahead and assign id.a values.

## unjoined.b.shp might include deletions.
unjoined.b1.shp <- sf::st_read('unjoined_b1.shp')

# unjoined.b1.deletions <-   unjoined.b1 %>%
#     dplyr::filter(!(id.b %in% unjoined.b1.shp$id_b)) %>%
#     dplyr::mutate(year2019=TRUE, year2021=FALSE,
#                   length_ft = units::drop_units(length_ft),
#                   WetRad_ft = units::drop_units(WetRad_ft))
## The deletions might need to be reshaped.
# sf::st_write(unjoined.b1.deletions, 'unjoined_b1_deletions.shp')
# unjoined.b1.deletions.shp <- sf::st_read('unjoined_b1_deletions.shp')
### None of the deletions are relevant, they are already digitized separately.

unjoined.b2 <- unjoined.b1.shp %>%
    dplyr::mutate(
      id.a=(1:n()) + max(zones.a$id.a), 
      year2009=F, year2011=F, year2013=F, year2015=F, year2017=F, year2018=F, 
      year2019=TRUE, year2021=TRUE) %>%
    dplyr::rename(id.b=id_b, County=CountyName)

```

```{r}
### unjoined.a.shp includes some pivots that were first spotted during this review (2022 imagery)
### there might be a few deletions here also. (year_2022 == FALSE)
unjoined.a1.shp <- sf::st_read('unjoined_a1.shp') %>%
    dplyr::mutate(
      year2019 = id_a %in% unjoined.a1$id.a, 
      year2021 = TRUE) %>%
    dplyr::rename(id.a = id_a)

unjoined.a1.deletions <-   unjoined.a1 %>%
    dplyr::filter(!(id.a %in% unjoined.a1.shp$id.a)) %>%
    dplyr::mutate(year2019 = FALSE, year2021 = FALSE)

unjoined.a1.new <- unjoined.a1.shp %>%
  dplyr::filter(is.na(id.a)) %>%
  dplyr::mutate(id.a = (1:n()) + max(unjoined.b2$id.a)) %>%
  dplyr::mutate(dplyr::across(
      year2009:year2018, ~F))

unjoined.a2 <- dplyr::bind_rows(
  unjoined.a1.shp %>%
    dplyr::filter(!is.na(id.a)),
  unjoined.a1.deletions,
  unjoined.a1.new,
  .id='set') %>%
  dplyr::mutate(
    dplyr::across(year2009:year2018, as.logical))

unjoined.a2 %>%
  dplyr::filter(scutils::duplicated2(id.a)) %>%
  View() ## zero entries
```





```{r}
### joined.a0.shp includes deletions.
joined.a0.shp <- sf::st_read('joined_a0.shp') %>%
  dplyr::mutate(year2019=T, year2021=T) %>%
  dplyr::rename(id.a=id_a)

joined.a1 <- dplyr::bind_rows(
  joined.a0.shp,
  joined.a0 %>%
    dplyr::filter(!(id.a %in% joined.a0.shp$id.a)) %>%
    dplyr::mutate(year2019=T, year2021=F)) %>%
  dplyr::mutate(
    dplyr::across(year2009:year2018, as.logical))
```


```{r}
### multi_joins.a1.shp has a few that were first spotted now, but mostly spotted in 2021 study.
### not sure there is any way to tell the difference, so figure all the new ones were all spotted in 2021.
### also a few deletions.
multi_joins.a1.shp <- sf::st_read('multi_joins_zone_1a.shp') %>%
  dplyr::rename(id.a=id_a) %>%
  dplyr::mutate(year2019=T, year2021=T)

multi_joins.a1.deletions <- multi_joins_zone1.a %>%
  dplyr::filter(!(id.a %in% multi_joins.a1.shp$id.a)) %>%
  dplyr::mutate(year2019=T, year2021=F)

multi_joins.a1.new <- multi_joins.a1.shp %>%
  dplyr::filter(is.na(id.a)) %>%
  dplyr::mutate(id.a = (1:n()) + max(unjoined.a2$id.a)) %>%
  dplyr::mutate(dplyr::across(
      year2009:year2018, ~F))

multi_joins.a2 <- dplyr::bind_rows(
  multi_joins.a1.shp %>%
    dplyr::filter(!is.na(id.a)), 
  multi_joins.a1.deletions,
  multi_joins.a1.new) %>%
  dplyr::mutate(
    dplyr::across(year2009:year2018, as.logical)) %>%
  dplyr::filter( !sf::st_is_empty(geometry)) ## there was a duplicate entry with empty geometry.
```


```{r}

zones2 <- dplyr::bind_rows(
  'unjoined.a' = unjoined.a2, 
  'joined.a' = joined.a1,
  'multi_joins.a' = multi_joins.a2,
  'unjoined.b' = unjoined.b2,
  .id='id') %>%
  dplyr::select(
    id, id.a, Type, Type2, County, 
    year2009, year2011, year2013, year2015, 
    year2017, year2018, year2019, year2021)

```

```{r}
## Test the id values.
sum(duplicated(zones2$id.a))

# sum(duplicated(multi_joins.a2$id.a))

```


```{r}
## make valid
zones3 <- zones2 %>%
  sf::st_make_valid(zones2) %>%
  dplyr::select(-id)


## inspect intersections.
zones3.intersections <- sf::st_intersection(zones3) %>%
  dplyr::filter(n.overlaps > 1) %>%
  sf::st_collection_extract('POLYGON') %>%
  dplyr::select(-origins)

##### I edited this shapefile, so don't overwrite it.
## sf::st_write(zones3, 'zones3.shp')

#### I didnt edit this one, I used it to iterate through.
## sf::st_write(zones3.intersections, 'zones3_intersections.shp')

```

```{r}
## read in the edited shapefile.
zones3.shp <- sf::st_read('zones3.shp') 

## calculate polygon areas (acres)
zones4 <- zones3.shp %>%
  dplyr::mutate(Acres = units::drop_units(sf::st_area(geometry))/4046.86)

## There are some little split ends...
# length(unique(zones3.shp$id_a)) ## 3496
zones4 %>%
  dplyr::filter(duplicated(id_a))
## 2750 once and 2786 twice
zones4 %>%
  dplyr::filter(Acres < 1)
## 2750 once and 2786 twice. Perfect.

## filter out negligible areas, which are also duplicates.
zones5 <- zones4 %>%
  dplyr::filter(Acres > 0.1)
# length(unique(zones5$id_a)) ## 3496, perfect

## split multi-polygons
# sf::st_geometry_type(zones3.shp) %>% unique()
## they are all just polygons, no multipolygons


## test that any remaining intersections are minimal in area
zones5.intersections <- sf::st_intersection(zones5) %>%
  dplyr::filter(n.overlaps > 1) %>%
  sf::st_collection_extract('POLYGON') %>%
  dplyr::select(-origins) %>%
  dplyr::mutate(Acres = units::drop_units(sf::st_area(geometry))/4046.86)


## Remove all of the little intersecting pieces?
## sf::st_disjoint (?) idk, I don't need to deal with this right now.
## sf::st_difference() or sf::st_sym_difference()

```


```{r}

## assign county values.
counties <- scsf::counties

sf::st_crs(counties) 

sf::st_crs(zones5)

zone.midpts <- zones5 %>%
  sf::st_transform(2273) %>%
  sf::st_centroid() %>%
  sf::st_join(
    sf::st_transform(counties, 2273))

county.mixup <- zone.midpts %>%
  dplyr::filter(!is.na(County.x) &
                  County.x != County.y) 

# sf::st_write(county.mixup, 'county_mixup.shp')
### Reviewed in QGIS, each is very close to a county line.

county.missing <- zone.midpts %>%
  dplyr::filter(is.na(County.y))

# sf::st_write(county.missing, 'county_missing.shp')
### Reviewed in QGIS, mostly in GA, which I accidentally digitized late at night.
## a few in NC just across the line near Landrum,SC and Tryon, NC. North Pacolet River.

county.join <- zone.midpts %>%
  sf::st_drop_geometry() %>%
  dplyr::select(id_a, County = County.y)

zones6 <- zones5 %>%
  dplyr::select(-County) %>%
  dplyr::left_join(county.join, 'id_a') %>%
  dplyr::filter(!is.na(County))
```


```{r}
## is Type2 needed?
dplyr::filter(zones6, Type != Type2)
## 0 rows. forget about it.

unique(zones6$Type)

sum(is.na(zones6$Type))
## assign Type values.

zones7 <- zones6 %>%
  dplyr::select(-Type2) %>%
  dplyr::mutate(Type = dplyr::if_else(is.na(Type), "Center Pivot", Type))
```


```{r}
zones <- zones7

usethis::use_data(zones, overwrite=T)
```
